---
output:
  html_document:
    css: styles.css
---

<!-- Optional: include external CSS for fonts -->
<link href="https://fonts.googleapis.com/css2?family=Roboto:wght@400;700&display=swap" rel="stylesheet">

<style>
body {
    background-image: url('https://xmple.com/wallpaper/white-gradient-blue-linear-3840x2160-c2-add8e6-ffffff-a-285-f-14.svg');
    background-size: cover;
    background-position: center;
    background-repeat: no-repeat;
    font-family: 'Roboto', sans-serif;
}
</style>


# How has Stephen Curry's individual performance evolved throughout various NBA regular seasons, and what impact has it had on his team's overall success? {.tabset}

Todd Ganbat   


<!-- ---------------------------------------------------------------------- -->
<!-- Data Details - Task 1 Tab -->
<!-- ---------------------------------------------------------------------- -->

## Data Details

#### Question 1 

How has Stephen Curry's performance in NBA regular seasons changed over time, and how does his individual performance influence his team's success?

#### Question 2 

__Description of data's origin__:

The dataset provided originates from Basketball-Reference.com, a reputable source for basketball-related statistics and information. It was curated specifically for educational purposes, with the intention of facilitating learning and analysis within academic or instructional contexts.

This dataset encompasses a comprehensive range of statistics pertaining to the Golden State Warriors, a prominent team in the National Basketball Association (NBA), as well as detailed metrics regarding the individual performance of Stephen Curry, a distinguished player renowned for his exceptional skills and contributions to the sport. The data spans a substantial timeframe, covering the years from 2009 to 2024, offering a rich and extensive resource for studying the performance trends and dynamics within the NBA during this period.

It's important to note that this dataset is not intended for commercial or operational use, but rather for educational exploration and academic inquiry. As such, it should be utilized with a focus on learning, analysis, and research within appropriate educational or instructional settings.

__Description of the files included with the data__:

- Meta Data: The Description for each observation in the data set.
- Read me: Text file that has information about license, data type (census, survey, etc.), date last updated, data limitations, data privacy or other ethical considerations.
- Rmd file: The document that has the code and explanatory text about the dataset and its analysis.
- R project file: R studio workplace serves as a container for this project's code and explanatory text.

__Description of Data Privacy and Assumptions__:

The dataset sourced from Basketball-Reference.com serves as a valuable resource for educational purposes, facilitating in-depth analysis and exploration within academic or instructional contexts.

Privacy Guidelines & Ethical Considerations:

Educational Analysis: The data is intended for educational analysis, allowing users to delve into various aspects of basketball statistics and performance trends.

Non-Commercial Use: It is not to be used for commercial purposes or operational activities but rather for academic exploration and research.

Ethical Considerations: Users are expected to adhere to ethical standards and guidelines, ensuring the responsible and ethical use of the data.

Attribution: Proper attribution to Basketball-Reference.com should be provided whenever the dataset is utilized or referenced in academic or research endeavors.

Assumptions:

Accuracy and Reliability: The dataset is assumed to be accurate and reliable, sourced from a reputable platform known for providing comprehensive and meticulously curated basketball statistics.

Scope and Coverage: The data covers statistics related to the Golden State Warriors and Stephen Curry spanning from 2009 to 2024, encompassing a diverse range of performance metrics.

Educational Intent: The dataset is assumed to be compiled with the specific purpose of supporting educational initiatives, enabling students and researchers to analyze and interpret basketball-related data for learning and research purposes.

__Data Limitations__:

While the dataset sourced from Basketball-Reference.com offers valuable insights into basketball statistics, it is essential to acknowledge certain limitations that may impact its usability and interpretation:

Incomplete Coverage: The dataset may not capture every aspect of the Golden State Warriors' performance or Stephen Curry's individual contributions during the specified time period. Certain events, games, or statistical nuances may not be fully represented, leading to potential gaps in the analysis.

Data Quality: While efforts are made to ensure the accuracy and reliability of the data, occasional discrepancies or errors may exist. Users should exercise caution when interpreting the statistics and cross-reference with other reliable sources whenever possible.

Scope of Metrics: The dataset primarily focuses on traditional basketball statistics such as points, assists, rebounds, and shooting percentages. It may lack coverage of more advanced metrics or contextual factors that could provide a deeper understanding of player performance and team dynamics.

Subjectivity in Interpretation: Interpretation of the data is subjective and may vary based on individual perspectives or analytical methodologies. Users should approach the analysis with an awareness of potential biases and exercise critical judgment in drawing conclusions.

Temporal Considerations: The dataset covers a specific time frame from 2009 to 2024, and trends or patterns observed within this period may not necessarily reflect long-term or future developments in basketball or player performance.

External Factors: The performance of the Golden State Warriors and Stephen Curry may be influenced by various external factors such as injuries, coaching changes, opponent strength, and rule changes, which may not be fully captured or accounted for in the dataset.

__Data Type__:

The data sourced from Basketball-Reference.com can be classified as sports statistics data, which is a subset of observational data. This type of data is collected through observation or measurement of natural phenomena, in this case, basketball games and player performances. It differs from traditional survey or census data, which involve systematic data collection through structured questionnaires or population counts.

#### Question 3 

I utilized a dataset consisting of:

* Stephen Curry's regular season statistics from 2009-2010 season to 2023-2024 season (points, assists, rebounds, steals, blocks, turnovers, shooting percentages, etc.) [here](https://www.basketball-reference.com/players/c/curryst01.html).

* The win/loss ratio of the Golden State Warriors spanning from the 2009-2010 season to the 2023-2024 season [here](https://www.basketball-reference.com/teams/GSW/).

I gathered the data for this task by extracting it from Basketball-Reference.com. Here's how I obtained the data:

```{r echo = TRUE, message = FALSE, warning = FALSE}
library(rvest)
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)

curry_url <- "https://www.basketball-reference.com/players/c/curryst01.html"

curry_data <- read_html(curry_url)

table_data <- curry_data %>%
  html_table(header = TRUE)

clean_data1 <- table_data[[2]]

clean_data1 <- clean_data1[-nrow(clean_data1), ]
colnames(clean_data1)[colnames(clean_data1) == "Season"] <- "Year"
```

Here's how I retrieved the data for the Golden State Warriors:

```{r echo = TRUE, message = FALSE, warning = FALSE}
gsw_url <- "https://www.basketball-reference.com/teams/GSW/"

gsw_data <- read_html(gsw_url)

table_data2 <- gsw_data %>%
  html_table(header = TRUE)

clean_data2 <- table_data2[[1]]

clean_data2 <- clean_data2 %>% select(Season, W, L, 'W/L%') 
clean_data2 <- clean_data2[order(clean_data2$Season), ]
clean_data2 <- clean_data2[64:78,]
```

Here's how I created a function to correctly merge two tables:

```{r echo = TRUE, message = FALSE, warning = FALSE}
colnames(clean_data2)[colnames(clean_data2) == "Season"] <- "Year"

clean_data1$Year <- as.character(clean_data1$Year)
clean_data2$Year <- as.character(clean_data2$Year)

clean_data_full <- full_join(clean_data1, clean_data2, by = "Year")
clean_data_full <- clean_data_full[1:15,]

colnames(clean_data_full)[colnames(clean_data_full) == "3P%"] <- "Threepoint_per"

colnames(clean_data_full)[colnames(clean_data_full) == "W/L%"] <- "Winloss_per"

colnames(clean_data_full)[colnames(clean_data_full) == "FG%"] <- "Fieldgoal_per"

colnames(clean_data_full)[colnames(clean_data_full) == "eFG%"] <- "Eff_fg_per"
```

I have chosen not to store the data locally since it's small and can be easily downloaded as long as an internet connection is available.

#### References

__rvest__: Wickham H (2024). _rvest: Easily Harvest (Scrape) Web Pages_. R package version 1.0.4,
  <https://CRAN.R-project.org/package=rvest>.
  
__tidyverse__:  Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A,      Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP,    Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H
   (2019). “Welcome to the tidyverse.” _Journal of Open Source Software_, *4*(43), 1686.                   doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
   
__stringr__: Wickham H (2023). _stringr: Simple, Consistent Wrappers for Common String Operations_. R      package version 1.5.1, <https://CRAN.R-project.org/package=stringr>.

__dplyr__: Wickham H, François R, Henry L, Müller K, Vaughan D (2023). _dplyr: A Grammar of Data           Manipulation_. R package version 1.1.4, <https://CRAN.R-project.org/package=dplyr>.

__ggplot2__: H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

#### Additional Files

To access information regarding the data's license, type (such as census or survey), date of last update, any limitations associated with the data, as well as considerations for data privacy, please consult the __"read_me.txt"__ file and the __"meta_data.csv"__ file. These documents provide detailed documentation and guidance on these important aspects of the dataset.

<!-- ---------------------------------------------------------------------- -->
<!-- Blog post - Task 2 Tab -->
<!-- ---------------------------------------------------------------------- -->

## Blog post 

### Who is Stephen Curry? 

Stephen Curry is a professional basketball player known for his exceptional shooting skills. And he happens to be one of my favorite athletes! Curry plays as a point guard for the Golden State Warriors in the NBA. He is renowned for revolutionizing the game with his extraordinary three-point shooting ability, consistently making shots from long distances with remarkable accuracy. His impressive ball-handling and playmaking skills also make him a standout player.

Curry has won multiple NBA championships and has been named the league's Most Valuable Player (MVP) multiple times. Off the court, he is known for his philanthropic efforts and engaging personality. His influence extends beyond basketball, making him a global sports icon and a source of inspiration for aspiring athletes. You can find more about his career and achievements [here](https://en.wikipedia.org/wiki/Stephen_Curry).

In his career, Curry has set numerous records and continues to be a dominant force in the NBA. His popularity has led to widespread recognition, and he has become a key figure in promoting the sport of basketball around the world.

### Why analyse Stephen Curry's stats? 

Many analysts are motivated by their passion for basketball and dedication to understanding the game's intricacies. For many, analyzing Stephen Curry's stats is a privilege and a fascinating challenge. Some analysts focus on specific aspects, such as his shooting percentages from various distances. The pool of analysts has also become far more diverse as interest in advanced basketball metrics has grown and Curry's popularity has soared. 

Since Stephen Curry is the unanimously greatest shooter the basketball world has ever seen, I'm curious about understanding how Curry's performance and his offensive skillset impact the overall success of the Golden State Warriors. 

### Where can we get necessary data of Stephen Curry and his team? 

The data for analyzing Stephen Curry's stats can be gathered from Basketball-Reference.com. The results are as follows.

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)

clean_data_full |>
  kbl() |>
  kable_styling()
```

### What do the data reveal about Stephen Curry's individual performance and its impact on his team's success?

I chose certain questions about Stephen Curry's Points Per Game Over the Years, relationship between Stephen Curry's Points Per Game and Team Win-Loss Percentage, relationship between Stephen Curry's Points Per Game and Three Point Attempts. Analyzing such statistics can provide valuable insights for several reasons as it displayed below.

__Performance Trends and Impact__:

By analyzing Curry's points per game (PPG) over the years, you can discern patterns in his performance. This analysis sheds light on his development as a player, his ability to adapt his game over time, and his consistency. Additionally, it reveals how his scoring averages have correlated with the number of wins each season.

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)

ggplot(clean_data_full, aes(x = Year, y = PTS, label = W, color = W, group = 1)) +
  geom_line() +                  # Line plot
  geom_point() +                 # Points
  geom_text(nudge_y = 1) +       # Labels, nudge_y adjusts the vertical position
  scale_color_gradient(low = "red", high = "green") + # Color gradient for wins
  labs(title = "Stephen Curry's Points Per Game Over the Years",
       x = "Year",
       y = "Points Per Game",
       color = "Wins") +
  theme_bw()   


```

This result demonstrates the progression of Stephen Curry's offensive abilities over the years and the related annual team wins. It emphasizes the ascent of his scoring performance to elite levels. While this enhancement has greatly contributed to his personal success on the basketball court, it has also translated into more team wins in most years. Other than that, we can also see there are two exceptional outliers during the 2011-12 and 2019-2020 seasons, attributable to Curry's injuries during those periods.

__Impact on Team Success__:

This analysis can reveal how Curry's scoring influences the overall success of his team.

```{r echo = FALSE, message = FALSE, warning = FALSE}
correlation <- cor(clean_data_full$Winloss_per, clean_data_full$PTS)

ggplot(clean_data_full, aes(x = Winloss_per, y = PTS)) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "green") +  # Add correlation line
  labs(title = "Stephen Curry's Points Per Game vs Team Win-Loss Percentage",
       x = "Win-Loss Percentage",
       y = "Points Per Game") +
  annotate("text", x = max(clean_data_full$Winloss_per), y = min(clean_data_full$PTS), 
           label = paste("Correlation coefficient:", round(correlation, 2)), hjust = 1, vjust = 0) +
  theme_bw()

```

This outcome highlights a significant correlation, suggesting that his scoring plays a pivotal role in the team's successes. Understanding this relationship can be invaluable for devising game strategies and gaining insights into his role within the team. Additionally, there are couple of outliers in this dataset, attributable to either shortened seasons or periods of Curry's injury.

__Shooting Efficiency__:

By analyzing the relationship between his PPG and his three-point attempts, you can assess his shooting efficiency and the impact of his three-point shooting on his overall scoring. This can provide insights into his scoring methods and highlight the importance of his elite three-point shooting ability in his scoring arsenal.

```{r echo = FALSE, message = FALSE, warning = FALSE}
ggplot(clean_data_full, aes(x = `3PA`, y = PTS)) +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "green") +  # Add correlation line
  labs(title = "Stephen Curry's Points Per Game vs Three Point Attempts",
       x = "Three Point Attempts",
       y = "Points Per Game") +
  annotate("text", x = max(clean_data_full$`3PA`), y = min(clean_data_full$PTS), 
           label = paste("Correlation coefficient:", round(cor(clean_data_full$`3PA`, clean_data_full$PTS), 2)), hjust = 1, vjust = 0) +
    theme_bw()
```

This outcome demonstrates a clear relationship: as Stephen Curry's three-point attempts rise, his scoring average directly follows suit. Building upon our earlier observation of how his scoring average impacts his team's success, this current insight underscores the pivotal role of three-point attempts in shaping his scoring average. The correlation coefficient of 0.9 solidifies this conclusion, indicating a strong positive correlation between these variables. Additionally, there are couple of outliers in this dataset, attributable to either shortened seasons or periods of Curry's injury.

### What else could we learn from the analysis?

While examining regular stat lines such as points, shooting percentages, and wins is valuable, delving into advanced statistics like box plus-minus and usage rate can provide deeper insights. I'd like to analyze Stephen Curry's impact not only on the Golden State Warriors but also on the entire NBA. Specifically, I'd explore:

* The increase in three-point shot attempts by NBA players since Curry entered the league.
* The overall performance of the Golden State Warriors without Stephen Curry.
* A comparison of Curry's box plus-minus and usage rate with those of all-time great players throughout NBA history.

### References

__Basketball-Reference.com__ (2024). _Stephen Curry_.  <https://www.basketball-reference.com/players/c/curryst01.html>.

__Basketball-Reference.com__ (2024). _Golden State Warriors_.  <https://www.basketball-reference.com/players/c/curryst01.html>.

__Zhu H__ (2024). _kableExtra: Construct Complex Table with 'kable' and Pipe Syntax_. R package
  version 1.4.0, <https://CRAN.R-project.org/package=kableExtra>.

<!-- ---------------------------------------------------------------------- -->
<!-- Behind the Scenes - Task 3 Tab -->
<!-- ---------------------------------------------------------------------- -->

## Behind the Scenes

#### Question 4

*Tell us about a part(s) of your data processing or analysis that weren't "sexy" and wouldn't typically be included in a blog post. (e.g. Was their any data drudgery or time intensive wrangling? Were there any repetitive tasks or manual tasks? If it was easy, describe what made it easy?)*

Consolidating the data into a uniform format was a challenging endeavor, but it significantly reduced the need for manual handling, streamlining the overall process. Furthermore, optimizing the presentation of each chart demanded a concerted effort, as ensuring clarity and effectiveness in visual communication required meticulous attention to detail and thoughtful design adjustments.

#### Question 5

*Were there any challenges that you faced in conducting this analysis. These may take the form of data limitations or coding challenges? (e.g. Was there anything in your analysis that you were not anticipating when you started? Did you have to change your intended scope?)*

Including certain types of individual statistics posed challenges, so I opted not to include them, as the primary focus was on Stephen Curry's individual performance and its impact on the team. While I was eager to delve into each type of statistic, doing so might have complicated the analysis or strayed from the main context. Therefore, I focused on key metrics such as overall team wins, win-loss percentage, points per game, and three-point percentage.

As for the time-intensive data wrangling, merging the team and player stat lines was not overly difficult. However, some rows required cleaning to ensure successful merging and preparation of the final dataset for analysis.

#### Questions 6 

*Tell us about any imperfect parts of your work or how would like to expand this analysis in future? (We don't expect you to solve everything in one assignment.)*

Beyond the conventional statistics like points scored, shooting accuracy, and team wins, there lies a wealth of insights in advanced metrics such as box plus-minus, usage rate and many more. These metrics offer a nuanced understanding of a player's impact on both individual games and overall team performance.

For my future analysis, I aim to delve into Stephen Curry's influence, not only within the Golden State Warriors but across the broader NBA landscape. Specifically, I intend to explore several key areas:

* The Evolution of Three-Point Shooting: Since Curry's entry into the league, there has been a noticeable uptick in three-point shot attempts by NBA players. By examining trends over time, I seek to uncover the extent of Curry's influence on the league's playing style and strategic emphasis on long-range shooting.

* Warriors' Performance Sans Curry: Beyond his individual contributions, I will scrutinize the Golden State Warriors' performance in games where Curry was absent. This analysis will shed light on Curry's importance to the team's success and highlight any discernible differences in their gameplay and outcomes in his absence.

* Comparative Analysis of Advanced Metrics: To contextualize Curry's impact within the pantheon of basketball greats, I plan to compare his box plus-minus and usage rate with those of historical NBA legends. By comparing Curry's metrics against players renowned for their dominance and influence, we can better appreciate his unique standing in the annals of basketball history.



